Manuscript

### 2019-01-10 manuscript.Rmd
update manuscript, fix table, venndiagrams

floatsintext      : no
figuresintext     : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(papaja)
# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# update wordcount in Addins
```

@@

## Comparison of G2P drug-gene associations and CCLE, CTRP, and GDSC

```{r functions, echo=FALSE, include=FALSE}
e2exp_converter <- function(string) {
  e <- formatC(as.numeric(string), format = "e", digits = 0)
  exp <- sub("(\\d)e-0(\\d)", "$p_k < \\1^{-\\2}$", e, perl = TRUE)
  return(exp)
}
```

@@

# Wilcoxon test results
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_exp <- e2exp_converter(ccle_signif_g2p$FDR_cutoff[1])
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ccle_signif_g2p_grid_exp <- e2exp_converter(ccle_signif_g2p_grid$FDR_cutoff[1])
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_g2p_lineage_exp <- e2exp_converter(ccle_signif_g2p_lineage$FDR_cutoff[1])
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")
ccle_signif_grid_lineage_exp <- e2exp_converter(ccle_signif_grid_lineage$FDR_cutoff[1])

ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_exp <- e2exp_converter(ctrp_signif_g2p$FDR_cutoff[1])
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
ctrp_signif_g2p_grid_exp <- e2exp_converter(ctrp_signif_g2p_grid$FDR_cutoff[1])
ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_g2p_lineage_exp <- e2exp_converter(ctrp_signif_g2p_lineage$FDR_cutoff[1])
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")
ctrp_signif_grid_lineage_exp <- e2exp_converter(ctrp_signif_grid_lineage$FDR_cutoff[1])

gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_exp <- e2exp_converter(gdsc_signif_g2p$FDR_cutoff[1])
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
gdsc_signif_g2p_grid_exp <- e2exp_converter(gdsc_signif_g2p_grid$FDR_cutoff[1])
gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_g2p_lineage_exp <- e2exp_converter(gdsc_signif_g2p_lineage$FDR_cutoff[1])
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
gdsc_signif_grid_lineage_exp <- e2exp_converter(gdsc_signif_grid_lineage$FDR_cutoff[1])

@@

```{r table-g2p-nonCID, echo=FALSE, results='asis'}
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize("Compound identifier" = DrugID, "Number of G2P interpretations" = n, "Proportion of excluded interpretations (%)" = printnum(n / 402 * 100))
colnames(summ)[colnames(summ) == "Drug"] <- "Compound"
summ$Compound[summ$Compound == "egfr"] <- "EGFR"
summ$Compound[summ$Compound == "mab"] <- "mAb"

apa_table(summ, caption = "Distribution of level-A G2P drugs excluded from analyses",
          note = "mAb = monoclonal antibody; NA indicates G2P interpretation does not have an associated compound")
```

@@

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P genes for each G2P drug tested in the CCLE, CTRP, and GDSC drug screens (*Figures 1B and 2A*). At an unadjusted $\alpha = 0.05$, Wilcoxon rank-sum tests revealed that in CCLE, the AUC distributions of mutant and wildtype groups for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a false discovery rate (FDR) < 5%, selumetinib-*BRAF* ($p$ `r printp(ccle_signif_g2p_grid[ccle_signif_g2p_grid$Drug_Gene == "selumetinib_BRAF",]$p.adj)`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid)` significant combinations, one of which was selumetinib-*BRAF*.

@@

We computed Spearman correlation coefficients ($r_s$) between AUC of each drug-gene association and gene expression (RPKM) and copy number (log$_2$ ratio) of the gene in the association

@@

- What do these correlations with mutation status, gene expression, and copy number look like when you constrict to a specific lineage?
- When you restrict, do the correlations become more distinct?
- Are the lineage-specific results approximately the same as the pancancer results?
- Looking at a histogram of p-values, if the histogram is flat, then nothing is significant (note from meeting on 11/16)

@@

```{r crispr-data, echo = FALSE, warning = FALSE, message = FALSE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

# Wilcoxon test results
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_g2p_gene_exp <- e2exp_converter(crispr_signif_g2p_gene$FDR_cutoff[1])
crispr_signif_g2p_gene_sig <- filter(crispr_signif_g2p_gene, !is.na(p.signif.adj))

@@

@@

When we don’t account for lineage, Wilcoxon rank-sum tests (`r crispr_signif_g2p_gene_exp`, FDR < 5%) revealed that mutations in `r nrow(crispr_signif_g2p_gene_sig)` genes significantly correlated with CERES score: *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[1]`* ($p$ `r printp(crispr_signif_g2p_gene_sig$p.adj[1])`) and *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[2]`* ($p$ `r printp(crispr_signif_g2p_gene_sig$p.adj[2])`).

Gene expression (RPKM) of `r nrow(crispr_spearman_res_GE_sig)` out of `r nrow(crispr_spearman_res_GE)` (`r printnum(nrow(crispr_spearman_res_GE_sig) / nrow(crispr_spearman_res_GE) * 100)`%) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for `r nrow(crispr_spearman_res_CN_sig)` out of `r nrow(crispr_spearman_res_CN)` (`r printnum(nrow(crispr_spearman_res_CN_sig) / nrow(crispr_spearman_res_CN) * 100)`%) G2P genes.

For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are similar to the results from CCLE, CTRP, and GDSC (Table 2).

>>>>

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(papaja)
library(bookdown)
library(VennDiagram)
# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# update wordcount in Addins
```

@@

```{r functions, echo=FALSE, include=FALSE}
e2exp_converter <- function(string, p_type) {
  e <- formatC(as.numeric(string), format = "e", digits = 1)
  if(p_type == "pk") {
    if(grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-(\\d+)", "$p_k = \\1^{-\\2}$", e, perl = TRUE)
    }
    if(!grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-0(\\d)", "$p_k = \\1^{-\\2}$", e, perl = TRUE)
    }
  }
  if(p_type == "p") {
    if(grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-(\\d+)", "$p = \\1^{-\\2}$", e, perl = TRUE)
    }
    if(!grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-0(\\d)", "$p = \\1^{-\\2}$", e, perl = TRUE)
    }
  }
  if(p_type == "padj") {
    if(grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-(\\d+)", "$p_{adj} = \\1^{-\\2}$", e, perl = TRUE)
    }
    if(!grepl("(\\d+)e-([1-9]+)", e)) {
      exp <- sub("(.+)e-0(\\d)", "$p_{adj} = \\1^{-\\2}$", e, perl = TRUE)
    }
  }
  return(exp)
}
```

@@

# Wilcoxon test results
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_exp <- e2exp_converter(ccle_signif_g2p$FDR_cutoff[1], p_type = "pk")
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ccle_signif_g2p_grid_exp <- e2exp_converter(ccle_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_g2p_lineage_exp <- e2exp_converter(ccle_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")
ccle_signif_grid_lineage_exp <- e2exp_converter(ccle_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_exp <- e2exp_converter(ctrp_signif_g2p$FDR_cutoff[1], p_type = "pk")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
ctrp_signif_g2p_grid_exp <- e2exp_converter(ctrp_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_g2p_lineage_exp <- e2exp_converter(ctrp_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")
ctrp_signif_grid_lineage_exp <- e2exp_converter(ctrp_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_exp <- e2exp_converter(gdsc_signif_g2p$FDR_cutoff[1], p_type = "pk")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
gdsc_signif_g2p_grid_exp <- e2exp_converter(gdsc_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_g2p_lineage_exp <- e2exp_converter(gdsc_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
gdsc_signif_grid_lineage_exp <- e2exp_converter(gdsc_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

@@

```{r table-g2p-nonCID, results="asis"}
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize("CID" = DrugID, "N_G2P" = n, "%" = printnum(n / nrow(g2p_nonCID) * 100))
colnames(summ)[colnames(summ) == "Drug"] <- "Compound"
summ$Compound[summ$Compound == "egfr"] <- "EGFR"
summ$Compound[summ$Compound == "mab"] <- "mAb"
summ <- summ[order(summ$N_G2P),]

apa_table(summ, align = "c", font_size = "small", caption = "Level-A G2P interpretations excluded from analyses", note = "N_G2P (%) = number of associated G2P interpretations, % is out of column total; mAb = monoclonal antibody; NA indicates G2P interpretation does not have an associated compound.")
```

@@

**Describe why we compare all pairwise combos rather than filtering for just the drug-gene combos in G2P**

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all pairwise combinations of G2P genes and drugs tested in the CCLE, CTRP, and GDSC drug screens (*Figures 1B and 2A*). At an unadjusted $\alpha = 0.05$, Wilcoxon rank-sum tests revealed that in CCLE, AUC significantly correlated with mutation status for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a false discovery rate (FDR) < 5%, selumetinib-*BRAF* (`r e2exp_converter(ccle_signif_g2p_grid[ccle_signif_g2p_grid$Drug_Gene == "selumetinib_BRAF",]$p.adj, p_type = "padj")`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),])` significant combinations, one of which was selumetinib-*BRAF*. There were `r nrow(gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),])` significant G2P associations in the GDSC dataset (`r gdsc_signif_g2p_grid_exp`), Of all the significant correlations between mutation status and AUC across datasets, the only genes represented were *KRAS* and *BRAF* (Figure\ \@ref(fig:drug-venn)).

```{r drug-venn, echo = FALSE, message=FALSE, warning=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")

ccle_dgs <- as.character(ccle_signif_g2p_grid[!is.na(ccle_signif_g2p_grid$p.signif.adj),]$Drug_Gene)
ctrp_dgs <- as.character(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),]$Drug_Gene)
gdsc_dgs <- as.character(gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),]$Drug_Gene)

drug_venn <- venn.diagram(list("CCLE" = ccle_dgs, "CTRP" = ctrp_dgs, "GDSC" = gdsc_dgs), lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 50, 120), cat.dist = c(0.001, 0.02, 0.0001), cat.cex = 1.5, cex = 1, filename = NULL)
# in GDSC only
drug_venn[[7]]$label  <- paste(sort(setdiff(setdiff(gdsc_dgs, ctrp_dgs), ccle_dgs)), collapse = "\n")
# in GDSC and CTRP only
drug_venn[[8]]$label <- paste(intersect(gdsc_dgs, ctrp_dgs), collapse = "\n")
# in CTRP only
drug_venn[[9]]$label <- paste(sort(setdiff(setdiff(ctrp_dgs, gdsc_dgs), ccle_dgs)), collapse = "\n")
# in CTRP and CCLE only
drug_venn[[10]]$label <- intersect(ccle_dgs, ctrp_dgs)

grid.newpage()
grid.draw(drug_venn)
```

@@

We computed Spearman correlation coefficients ($r_s$) between AUC of each drug-gene association and gene expression (RPKM) and copy number (log$_2$ ratio) of the gene in the association...

@@

- What do these correlations with mutation status, gene expression, and copy number look like when you constrict to a specific lineage?
- When you restrict, do the correlations become more distinct?
- Are the lineage-specific results approximately the same as the pancancer results?
- Looking at a histogram of p-values, if the histogram is flat, then nothing is significant (note from meeting on 11/16)
- KS test results

@@

```{r crispr-data, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

@@

# Wilcoxon test results
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_g2p_gene_exp <- e2exp_converter(crispr_signif_g2p_gene$FDR_cutoff[1], p_type = "pk")
crispr_signif_g2p_gene_sig <- filter(crispr_signif_g2p_gene, !is.na(p.signif.adj))

@@

```{r crispr-venn, echo = FALSE}
crispr_venn <- venn.diagram(list("Gene Expression" = as.character(crispr_spearman_res_GE_sig$Hugo_Symbol), "Copy Number" = as.character(crispr_spearman_res_CN_sig$Hugo_Symbol)), fill = c("darkorchid4", "mediumorchid1"), lty = rep("blank", 2), alpha = c(0.5, 0.5), cat.pos = c(-2, 5), cat.cex = 1.5, cex = 1, filename = NULL)

# in GE only
crispr_venn[[5]]$label  <- paste(sort(setdiff(as.character(crispr_spearman_res_GE_sig$Hugo_Symbol), as.character(crispr_spearman_res_CN_sig$Hugo_Symbol))), collapse = "\n")
# in CN only
crispr_venn[[6]]$label <- paste(sort(setdiff(as.character(crispr_spearman_res_CN_sig$Hugo_Symbol), as.character(crispr_spearman_res_GE_sig$Hugo_Symbol))), collapse = "\n")
# intesection
crispr_venn[[7]]$label <- paste(sort(intersect(as.character(crispr_spearman_res_CN_sig$Hugo_Symbol), as.character(crispr_spearman_res_GE_sig$Hugo_Symbol))), collapse = "\n")

grid.newpage()
grid.draw(crispr_venn)
```

@@

When we don’t account for lineage, Wilcoxon rank-sum tests (`r crispr_signif_g2p_gene_exp`, FDR < 5%) revealed that mutations in `r nrow(crispr_signif_g2p_gene_sig)` genes significantly correlated with CERES score: *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[1]`* (`r e2exp_converter(crispr_signif_g2p_gene_sig$p.adj[1], p_type = "padj")`) and *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[2]`* (`r e2exp_converter(crispr_signif_g2p_gene_sig$p.adj[2], p_type = "padj")`). Gene expression (RPKM) of `r nrow(crispr_spearman_res_GE_sig)` out of `r nrow(crispr_spearman_res_GE)` (`r printnum(nrow(crispr_spearman_res_GE_sig) / nrow(crispr_spearman_res_GE) * 100)`%) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for `r nrow(crispr_spearman_res_CN_sig)` out of `r nrow(crispr_spearman_res_CN)` (`r printnum(nrow(crispr_spearman_res_CN_sig) / nrow(crispr_spearman_res_CN) * 100)`%) G2P genes (Figure\ \@ref(fig:crispr-venn)).

For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are more comprehensive than the results from CCLE, CTRP, and GDSC (Table 2).
